---
title: "Implementing MAPS in gene list recognition"
date: "`r format(Sys.Date(),'%e %B, %Y')`"
output:
  html_document:
    code_folding: hide
  pdf_document: default
---

```{r global, include=FALSE}
## Global options
knitr::opts_chunk$set(
  cache = FALSE,
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
```


```{r setup}
# Dependencies ------------------------------------------------------------
library(tidyverse)
library(ggplot2)
library(pROC)
library(caret)
library(ggplot2)
library(broom)
library(knitr)
library(kableExtra)

# Setup -------------------------------------------------------------------
maps_global <- read_csv('/Users/adrian/BroadIS/MAPS/01_data/maps_clean_global_Per_gene.csv') %>% 
  rename(gene_id = gene_ids)

metrics <- read_delim('/Users/adrian/BroadIS/MAPS/01_data/gnomad.v2.1.1.lof_metrics.by_gene.txt')

haploinsufficient_1  <- read_csv('/Users/adrian/BroadIS/MAPS/01_data/haploinsufficiency_severe_curated_2016.tsv',
                  col_names = c('gene'))
haploinsufficient_2  <- read_csv('/Users/adrian/BroadIS/MAPS/01_data/haploinsufficiency_moderate_curated_2016.tsv',
                  col_names = c('gene'))
haploinsufficient_3  <- read_csv('/Users/adrian/BroadIS/MAPS/01_data/haploinsufficiency_severe_curated_2016.tsv',
                  col_names = c('gene'))

haploinsufficient_final <- union(union(haploinsufficient_1$gene,
                                       haploinsufficient_2$gene), haploinsufficient_3$gene) %>% 
  tibble() %>% 
  rename(gene = ".") %>% 
  mutate(haploinsufficient = 1)
  
olfactory <- read_csv('/Users/adrian/BroadIS/MAPS/01_data/olfactory_receptors.tsv',
                                 col_names = c('gene')) %>% 
  mutate(olfactory = 1)

```

# {.tabset}

## 1) Haploinsufficiency

<h1 align="center"> HC </h2>

```{r}
# Join the tables
maps_haplo <- maps_global %>% 
  left_join(metrics) %>% 
  left_join(haploinsufficient_final) %>% 
  mutate(haploinsufficient = factor(ifelse(is.na(haploinsufficient), 0, haploinsufficient)))
```

#### *haploinsufficincy ~ MAPS*

```{r}
# Logit for just MAPS
haploinsufficient_MAPS_logit <- 
  glm(haploinsufficient ~ MAPS, 
      data = maps_haplo %>% filter(lof_csq_collapsed == 'HC'), 
      family = "binomial")

kable(tidy(haploinsufficient_MAPS_logit), digits = 3) %>%
  kable_styling(full_width = FALSE)
```

#### *haploinsufficincy ~ oe_lof_upper*

```{r}
# Logit for just oe_lof_upper
haploinsufficient_lof_logit <- 
  glm(haploinsufficient ~ oe_lof_upper, 
      data = maps_haplo %>% filter(lof_csq_collapsed == 'HC'), 
      family = "binomial")

kable(tidy(haploinsufficient_lof_logit), digits = 3) %>%
  kable_styling(full_width = FALSE)
```

#### *haploinsufficincy ~ oe_lof_upper + MAPS*

```{r}
# Add effect of MAPS
haploinsufficient_lof_maps_logit <- 
  glm(haploinsufficient ~ oe_lof_upper + MAPS, 
      data = maps_haplo %>% filter(lof_csq_collapsed == 'HC'), 
      family = "binomial")

kable(tidy(haploinsufficient_lof_maps_logit), digits = 3) %>%
  kable_styling(full_width = FALSE)
```

***

<h1 align="center"> LC </h2>

#### *haploinsufficincy ~ MAPS*

```{r}
# Logit for just MAPS
haploinsufficient_MAPS_logit <- 
  glm(haploinsufficient ~ MAPS, 
      data = maps_haplo %>% filter(lof_csq_collapsed == 'LC'), 
      family = "binomial")

kable(tidy(haploinsufficient_MAPS_logit), digits = 3) %>%
  kable_styling(full_width = FALSE)
```

#### *haploinsufficincy ~ oe_lof_upper*

```{r}
# Logit for just oe_lof_upper
haploinsufficient_lof_logit <- 
  glm(haploinsufficient ~ oe_lof_upper, 
      data = maps_haplo %>% filter(lof_csq_collapsed == 'LC'), 
      family = "binomial")

kable(tidy(haploinsufficient_lof_logit), digits = 3) %>%
  kable_styling(full_width = FALSE)
```

####  *haploinsufficincy ~ oe_lof_upper + MAPS*

```{r}
# Add effect of MAPS
haploinsufficient_lof_maps_logit <- 
  glm(haploinsufficient ~ oe_lof_upper + MAPS, 
      data = maps_haplo %>% filter(lof_csq_collapsed == 'LC'), 
      family = "binomial")

kable(tidy(haploinsufficient_lof_maps_logit), digits = 3) %>%
  kable_styling(full_width = FALSE)
```
 
***
 
<h1 align="center"> Synonymous </h2>

#### *haploinsufficincy ~ MAPS*

```{r}
# Logit for just MAPS
haploinsufficient_MAPS_logit <- 
  glm(haploinsufficient ~ MAPS, 
      data = maps_haplo %>% filter(lof_csq_collapsed == 'synonymous_variant'), 
      family = "binomial")

kable(tidy(haploinsufficient_MAPS_logit), digits = 3) %>%
  kable_styling(full_width = FALSE)
```

#### *haploinsufficincy ~ oe_lof_upper*

```{r}
# Logit for just oe_lof_upper
haploinsufficient_lof_logit <- 
  glm(haploinsufficient ~ oe_lof_upper, 
      data = maps_haplo %>% filter(lof_csq_collapsed == 'synonymous_variant'), 
      family = "binomial")

kable(tidy(haploinsufficient_lof_logit), digits = 3) %>%
  kable_styling(full_width = FALSE)
```

#### Sanity check for missense as it has the same genes

```{r}
# Logit for just oe_lof_upper
missense_lof <- 
  glm(haploinsufficient ~ oe_lof_upper, 
      data = maps_haplo %>% filter(lof_csq_collapsed == 'missense_variant'), 
      family = "binomial")

kable(tidy(missense_lof), digits = 3) %>%
  kable_styling(full_width = FALSE)
```

####  *haploinsufficincy ~ oe_lof_upper + MAPS*

```{r}
# Add effect of MAPS
haploinsufficient_lof_maps_logit <- 
  glm(haploinsufficient ~ oe_lof_upper + MAPS, 
      data = maps_haplo %>% filter(lof_csq_collapsed == 'synonymous_variant'), 
      family = "binomial")

kable(tidy(haploinsufficient_lof_maps_logit), digits = 3) %>%
  kable_styling(full_width = FALSE)
```


## 2) Olfactory

<h1 align="center"> HC </h2>

```{r}
# Join the tables
maps_olfactory <- maps_global %>% 
  left_join(metrics) %>% 
  left_join(olfactory) %>% 
  mutate(olfactory = factor(ifelse(is.na(olfactory), 0, olfactory)))
```

#### *olfactory ~ MAPS*

```{r}
# Logit for just MAPS
olfactory_MAPS_logit <- 
  glm(olfactory ~ MAPS, 
      data = maps_olfactory %>% filter(lof_csq_collapsed == 'HC'), 
      family = "binomial")

kable(tidy(olfactory_MAPS_logit), digits = 3) %>%
  kable_styling(full_width = FALSE)
```

#### *olfactory ~ oe_lof_upper*

```{r}
# Logit for just oe_lof_upper
olfactory_lof_logit <- 
  glm(olfactory ~ oe_lof_upper, 
      data = maps_olfactory %>% filter(lof_csq_collapsed == 'HC'), 
      family = "binomial")

kable(tidy(olfactory_lof_logit), digits = 3) %>%
  kable_styling(full_width = FALSE)
```

#### *olfactory ~ oe_lof_upper + MAPS*

```{r}
# Add effect of MAPS
olfactory_lof_maps_logit <- 
  glm(olfactory ~ oe_lof_upper + MAPS, 
      data = maps_olfactory %>% filter(lof_csq_collapsed == 'HC'), 
      family = "binomial")

kable(tidy(olfactory_lof_maps_logit), digits = 3) %>%
  kable_styling(full_width = FALSE)
```

***

<h1 align="center"> LC </h2>

#### *olfactory ~ MAPS*

```{r}
# Logit for just MAPS
olfactory_MAPS_logit <- 
  glm(olfactory ~ MAPS, 
      data = maps_olfactory %>% filter(lof_csq_collapsed == 'LC'), 
      family = "binomial")

kable(tidy(olfactory_MAPS_logit), digits = 3) %>%
  kable_styling(full_width = FALSE)
```

#### *olfactory ~ oe_lof_upper*

```{r}
# Logit for just oe_lof_upper
olfactory_lof_logit <- 
  glm(olfactory ~ oe_lof_upper, 
      data = maps_olfactory %>% filter(lof_csq_collapsed == 'LC'), 
      family = "binomial")

kable(tidy(olfactory_lof_logit), digits = 3) %>%
  kable_styling(full_width = FALSE)
```

####  *olfactory ~ oe_lof_upper + MAPS*

```{r}
# Add effect of MAPS
olfactory_lof_maps_logit <- 
  glm(olfactory ~ oe_lof_upper + MAPS, 
      data = maps_olfactory %>% filter(lof_csq_collapsed == 'LC'), 
      family = "binomial")

kable(tidy(olfactory_lof_maps_logit), digits = 3) %>%
  kable_styling(full_width = FALSE)
```
 
***
 
<h1 align="center"> Synonymous </h2>

#### *olfactory ~ MAPS*

```{r}
# Logit for just MAPS
olfactory_MAPS_logit <- 
  glm(olfactory ~ MAPS, 
      data = maps_olfactory %>% filter(lof_csq_collapsed == 'synonymous_variant'), 
      family = "binomial")

kable(tidy(olfactory_MAPS_logit), digits = 3) %>%
  kable_styling(full_width = FALSE)
```

#### *olfactory ~ oe_lof_upper*

```{r}
# Logit for just oe_lof_upper
olfactory_lof_logit <- 
  glm(olfactory ~ oe_lof_upper, 
      data = maps_olfactory %>% filter(lof_csq_collapsed == 'synonymous_variant'), 
      family = "binomial")

kable(tidy(olfactory_lof_logit), digits = 3) %>%
  kable_styling(full_width = FALSE)
```

####  *olfactory ~ oe_lof_upper + MAPS*

```{r}
# Add effect of MAPS
olfactory_lof_maps_logit <- 
  glm(olfactory ~ oe_lof_upper + MAPS, 
      data = maps_olfactory %>% filter(lof_csq_collapsed == 'synonymous_variant'), 
      family = "binomial")

kable(tidy(olfactory_lof_maps_logit), digits = 3) %>%
  kable_styling(full_width = FALSE)
```







